---
title: "Life expectancy"
author: "Duc-Quang Nguyen"
date: "21 April 2016"
output: html_document
---

## Data

[US census bureau International Data Base (2015)](http://www.census.gov/data/developers/data-sets/international-database.html)


```{r setup, include=FALSE}
library(idbr) # devtools::install_github('walkerke/idbr')
library(countrycode)
library(ggplot2)
library(animation)
library(dplyr)
library(swiTheme)


countries <- c("France", "Switzerland", "Japan", "China", "United States", "Brazil", "Russia", "Morocco", "India")
country.iso2 <- countrycode(countries, "country.name", "fips104")

#translation.file <- "input/Population structures past and forecast - Sheet1.csv"

downloadData <- F

years <- 1990:2050
apikey <- "901083ee90dbd111d13f5e270ca00b4abf654be6"
data.file <- paste0("input/", paste(range(years), collapse = "_"), "_", 
  paste(country.iso2, collapse = "_"), ".csv")


```

```{r getData, echo = F}
if(downloadData) {
  idb_api_key(apikey)
  
  data <- do.call(rbind, lapply(country.iso2, function(iso) {
    cat("\n", iso)
    
    male <- idb1(iso, years, sex = 'male') %>%
      mutate(POP = POP * -1, SEX = 'Male')
    female <- idb1(iso, years, sex = 'female') %>%
      mutate(SEX = 'Female')
    
    rbind(male, female) %>% 
      mutate(abs_pop = abs(POP)) %>% select(-AREA_KM2)
  }))
  write.csv(data, file = data.file, row.names = F)
} else {
  data <- read.csv(data.file)
}
# for each year, country & sex: compute the proportion for each age
data <- data %>% group_by(NAME, FIPS, time, SEX) %>% mutate(yearlySum = sum(abs_pop)) %>% ungroup()
data$abs_prop <- data$abs_pop / data$yearlySum
data$prop <- ifelse(data$SEX == "Male", data$abs_prop * -1, data$abs_prop)

# sort data by youngest mean age at the beginning
d.sort <- data %>% filter(time == min(data$time)) %>% 
  group_by(NAME) %>% 
  summarise(mean = mean(sum(as.numeric(AGE * abs_pop, na.rm = T)) / 
    sum(abs_pop, na.rm = T), na.rm = T)) %>% 
  ungroup() %>% arrange(mean)
data$NAME <- factor(data$NAME, levels = as.character(d.sort$NAME))

```

```{r plot, echo = F}

plotyear <- function(
  data, 
  y,
  base_family = "OpenSans-CondensedLight",
  subtitle = "Evolution de la pyramide des âges entre 1990 et 2050",
  caption =  "Source: US Census Bureau IDB; idbr R package | swissinfo.ch",
  age = "AGE",
  maleFemale = c("Hommes", "Femmes"),
  countries = c( "Maroc", "Brésil", "Chine", "Russie", "Etats-Unis", "France", "Japon", "Suisse"  )
) 
  {

  dd <- data
  levels(dd$NAME) <- countries
  
  ggplot(data = filter(dd, time == y), aes(x = AGE, y = prop, fill = SEX, width = 1)) +
    coord_flip() +
    # gender bars
    geom_area(data = filter(dd, SEX == "Female", time == y)) +
    geom_area(data = filter(dd, SEX == "Male", time == y)) +
    # add median age
    facet_wrap( ~ NAME, ncol = 2) + 
    swi_theme(
      base_size = 24, 
      base_family = base_family, 
      subtitle = base_family, 
      title_family = base_family) +
  scale_fill_manual(values = c('#663333', '#336666'),
      labels=c(maleFemale[2], maleFemale[1])) +
  scale_y_continuous(
    labels = c("3%", "2%", "1%", "0%", "1%", "2%", "3%"),
    limits = c(-max(data$abs_prop), max(data$abs_prop)),
    name = y) +
  scale_x_continuous(name = age) +
  theme(
    legend.position = "top", 
    legend.title = element_blank(),
    legend.key.height = unit(0.5, "lines"),
    legend.key.width = unit(3, "lines")) + 
  guides(fill = guide_legend(reverse = TRUE)) +
  labs(
    subtitle = subtitle,
    caption = paste0(caption, " | Duc-Quang Nguyen"),
    title = y) + 
  theme(
    plot.title = element_text(size = 44, hjust = 0.5, face = "bold"),
    axis.text = element_text(size = 13),
    axis.title.x = element_text(size = 44),
    plot.caption = element_text(size = 14)
  ) 
}


lang <- 'RU'
plotyear(
  data, 
  1990,
  base_family = txt["font", lang],
  subtitle = txt["main.title", lang],
  caption =  txt["source", lang],
  age = txt["age", lang],
  maleFemale = c(txt["male", lang], txt["female", lang]),
  countries = c(txt["MO", lang], txt["BR", lang], txt["CH", lang], 
    txt["RS", lang], txt["US", lang], txt["FR", lang], txt["JA", lang], txt["SZ", lang])
) 

# drop some languages
txt <- txt[ , !(colnames(txt) %in% exludeLang), drop = F]


for(lang in colnames(txt)) {
  
  output <- paste0("int_populationStrucutres_", lang, ".gif")
  if(lang == "RU") {
   txt["main.title", lang] <- gsub('(.{1,75})(\\s|$)', '\\1\n', txt["main.title", lang])    
  }
  
  saveGIF({
	for(y in c(unique(data$time), rep(max(data$time), 3))) {
		print(plotyear(
		  data = data,
		  y = y,
		  base_family = txt["font", lang],
      subtitle = txt["main.title", lang],
      caption =  txt["source", lang],
      age = txt["age", lang],
      maleFemale = c(txt["male", lang], txt["female", lang]),
      countries = c(txt["MO", lang], txt["BR", lang], txt["CH", lang], 
        txt["RS", lang], txt["US", lang], txt["FR", lang], txt["JA", lang], txt["SZ", lang])
		  ))
	}
  }, movie.name = output, interval = 0.15, 
  ani.width = 700, ani.height = 860, loop = TRUE)
  
}








  
    

